Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-04-14 manuscript/Chapter2-manuscript.Rmd
Discussion finished (if everything OK)

---
title: "Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame"
output: 
  word_document:
    reference_docx: Chapter2-styles.docx
    fig_caption: yes
    pandoc_args: ["-Fpandoc-crossref"] # to have eq. numbers
bibliography: references2.bib
csl: chicago-author-date.csl
---

@@

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing in different foraging tactics or on different resources irrespective of sexual, morphological, or age-related differences [@Estes.etal2003;@Tinker.etal2008;@kernaleguenDeterminantsIndividualForaging2015;@Phillips.etal2017]. A growing body of evidence suggests that ecological interactions such as predator-prey interactions can drive this individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016]. When they hunt, predators often use techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their capacity to use them effectively is contingent on periods of extensive practice (i.e. experience). While hunting experience may be essential to maintain or improve foraging success, we have few empirical assessments of its role in predator foraging specialization [@Dukas2019] and its consequences on predator-prey interactions.

The development of foraging tactics is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. It is described as the relationship between success and cumulative experience, where extensive practice optimizes success when individuals reach an advanced level of experience. Empirical studies on human and non-human hunters show that experience optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. This body of work suggests that such optimization may thus promote individual foraging specialization (e.g. repeatedly using fast attacks) if the success of a hunting tactic is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that the costs/risks faced by hunters when switching foraging tactics may be offset through extensive practive, as individuals gain experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could result in both mechanisms operating at the same time within a predator population. For instance, some individuals may constantly encounter faster prey, which are proven to be more difficult to hunt in simulation and empirical studies [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. Thus, whether predators specialize or increase their flexibility with experience, learning to adjust their tactic to their prey should be essential to optimize their success.

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging [@Phillips.etal2017]. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specialization [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. This may in part be due to the limited information we have on the ecological contexts favouring specialization over flexibility in nature. However, in predator-prey systems, many studies have outlined that fluctuations in the predictability of prey encounters throughout a predator's lifetime may be a key factor shaping foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability (or resource-heterogeneity) hypothesis argues that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019]. When predator-prey interactions are directly observed, empirical evidence shows that predators often match, among other types of traits [see @Kishida.etal2006;@Hanifin.etal2008;@Brousseau.etal2018;@Reimche.etal2020], their locomotor and behavioural traits to those of their prey [@McGhee.etal2013;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022]. Yet, a recurring question that emerges from these studies is whether and how the behavioural adjustments of predators reflect individual differences in experience/learning [@Kelley.Magurran2011]. Thus, uncovering the role of learning in predator-prey systems where interactions are directly monitored would enable researchers to better predict the behavioural decisions and the success of predators when they are hunting.

@@

We tested whether foraging behaviour changed with experience, and compared the success of generalists vs specialists, using a multivariate, double-hierarchical generalized linear model (MDHGLM) [@leeDoubleHierarchicalGeneralized2006;@cleasbyQuantifyingPredictabilityBehaviour2015;@westneatBiologyHiddenResidual2014;@Mitchell.etal2016a;@ODea.etal2022]. We categorized the predators by their experience based on the number of matches they played (i.e. their cumulative experience). Since we monitored all predator players throughout their experience, they appeared in all of the three following experience categories. First, we assigned predators the status of "novice" when they had a cumulative experience below 100 matches. For example, a predator player that played 500 matches in total would be labelled as a novice in its first 100 matches. Following this logic, we labelled predators as "intermediate" hunters when they had a cumulative experience between 100 and 299 and matches. Lastly, we defined predators as "advanced" hunters when they had a cumulative experience equal or above 300, with a maximum of 499 cumulated matches.

@@

At the individual level, we detected differences in the mean movement speed among predators (i.e. the intercept of the mean speed for the predator ID; Table 3). These differences remained stable from the novice to the intermediate stages, and then increased at the advanced stage to reach a level higher than when predators were novice (Table 3 and Appendix S1: Figure S4). Predators displayed considerable differences in their degree of hunting specialization, and these differences increased with experience (i.e. the intercept of sigma for the predator ID increases with experience; Table 3 and Appendix S1: Figure S4). Thus, as individual predators gain experience, the population becomes more heterogeneous, with an assemblage of both specialist and flexible hunters. Moreover, some individual predators switched from a flexible to a specialized hunting tactic, and vice-versa, as they gained experience (Figure 2). Importantly, 13% of the predator population displayed a large change in foraging specialization/flexibility with experience (i.e. $\ge$0.2 change in standard deviation), 56% displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation), while 43% remained relatively stable (i.e. $\le$0.05 change in standard deviation).

@@

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We selected the individuals displayed in the figure by first substracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selecting those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization $\ge$0.2 standard deviations, while those on panel B have an increase in flexibility $\ge$"0.28 standard deviations.}

@@

Hunting experience is predicted to increase predator foraging specialization when resources are predictable [@Estes.etal2003;@Tinker.etal2008;@Tinker.etal2009]. However, when resources vary, others predict that learning with experience should increase behavioural flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013;@Wright.etal2022]. Our results show that there was no directional increase in either behavioural specialization or flexibility with experience in the population. Instead, both mechanisms operated simultaneously, such that each individual predator developed their own expertise and behavioural trajectory over successive encounters with prey; some specialized in fast-paced hunting, and others adopted a slower and more flexible tactic. The predators' tactics were matched to speed of the prey that they encountered, such that specialized cursorial hunters encountered similar groups of prey, while slower and flexible hunters encountered varying groups of prey, with no difference in success between the two types of hunters. This predator-prey behavioural matching grew stronger with predator experience, while the strenght of the relationship between the predator's hunting success and prey behaviour decreased. Together, our observations suggest that predators in *DBD* learn to adjust their hunting tactic to the behaviour of the prey that they encounter throughout their lifetime in the game. Moreover, the predator-prey behavioural escalation through experience suggests that learning may be a catalyst for arms races.

@@

Empirical research in humans and animals shows that task proficiency often increases nonlinearly with experience and stabilizes at an expert level [reviewed in @Dukas2019]. However, we cannot derive conclusions at the individual-level from these studies either because their models assume the same increase in expertise for each individual, or because it is unclear whether their data includes repeated measures of individual success through time. Our analyses revealed that there was no population increase in expertise over time; instead, individual predators in *DBD* displayed distinct patterns of expertise acquisition (Figure 1, Appendix S1: Figure S1A-S1B). We predicted that this would occur if, by chance, predators differed in the prey that they encountered, because some may be easier and others more difficult to capture. In nature, prey can use a variety of defenses such as physical armaments, toxins, camouflage, or antipredator behaviours to escape predation [@BrodieIII.BrodieJr.1999;@Bowen.etal2002;@brodieCoEvolutionPredatorsPrey2010;@Nomura.etal2011;@Carey.Wahl2011]. In *DBD* we previously showed that faster speed and space coverage is an effective antipredator strategy used by the prey [@FraserFranco.etal2022]. Other studies have also found that faster escapes are effective to escape predation [@Walker.etal2005;@Kelley.Magurran2011;Martin.etal2022]. Yet, despite that predators differed in the prey that they encountered, prey speed did not shape individual differences in expertise acquisition. It is however possible that other antipredator tactics mediated these individual responses. For example, it is well known in *DBD* that some prey groups use mobbing to confuse the predator or prevent it from capturing an injured individual. However, like natural predator-prey systems, direct predator-mobbing in *DBD* is dangerous and requires coordinated actions that are learned through experience [@Graw.Manser2007;@Dutour.etal2016;@Pitman.etal2017;@Dutour.etal2019;@Rampim.etal2020;@Szymkowiak2021], and was thus probably not a pervasive antipredator strategy. Other antipredator tactics known to confuse human and non-human predators during the attack sequence are unpredictable escapes and turning angles [@Jones.etal2011;@Briffa2013;@Herbert-Read.etal2017;@Moore.etal2017;@Richardson.etal2018]. While predators can adjust to unpredictable prey escapes [@Szopa-Comley.Ioannou2022], it could still be more arduous for a predator to learn and optimize its hunting success if it more often encountered prey using these tactics.

Alternatively, it is possible that individual variation in learning rates due to the time delay between hunting events drove the observed differences in expertise acquisition, irrespective of the prey groups that predators encountered. For example, in *DBD*, a predator that played 300 matches in the span of six months might forget more critical information (e.g. prey escape patterns or muscle memory) than one that played 300 matches in the span of 6 days. While this has not been formally tested, a greater delay between exposure events is hypothesized to attenuate or even negate the relationship between success and experience [@Endler1991;@Wright.etal2022]. Moreover, the predators in *DBD* may have differed in their capacity to process cues related to prey detection. For example, it is widely demonstrated that predators learn visual cues such as coloration to recognize and avoid unpalatable prey [@Skelhorn.etal2016;@Ward-Fear.etal2020;@McLellan.etal2021], or habitat characteristics to locate food patches [@Ehlinger1989;@Hughes.etal1992;@Warburton2003]. @Szopa-Comley.etal2020 also found consistent among individual variation in the capacity to detect cryptic and conspicuous prey in three-spined sticklebacks (*Gasterosteus aculeatus*). Thus, individual predators should differ in their expertise if they do not learn equally how to exploit the visual cues that facilitate prey detection [@dukasEvolutionaryBiologyAnimal2004;@Dukas2019;@Wright.etal2022].

@@

The predator population displayed an increase in foraging flexibility when reaching the intermediate level of experience. As it reached an advanced level of experience, it returned to a level of flexibility that was similar to novice levels. This is commonly observed in juvenile predators across the animal kingdom, because exploring and learning different tactics at this developmental stage is crucial to become a skillful hunter [@Vehanen2003;@Johnson.Wilbrecht2011;@Thiers.etal2014;@deGrissac.etal2016]. Thus, the predators in *DBD* were probably exploring and refining different tactics at this stage of experience. Interestingly, differences in individual foraging specialization among players increased with experience, which was related to the degree of variability in prey encounters. Predators that experienced predictable encounters with their prey across matches specialized in hunting at high speeds, while those that experienced less predictable encounters adopted a slower and flexible strategy. This is similar to other studies finding that fast-paced hunting is a highly specialized tactic suited for prey that use rapid escapes [@Endler1991;@Bro-Jorgensen2013;@Wilson.etal2018;@Szopa-Comley.Ioannou2022]. On the contrary, instead of focusing solely on one tactic, predators that experienced unpredictable encounters probably adjusted their behaviour to minimize the consequences of uncertainty.

ici une phrase qui rappelle que pas tout le monde a appris
 Thus, while individual differences in specialization did increase with experience, they were in part driven by a proportion of the population that displayed greater changes in behaviour compared to others.

As they gained experience, predators increasingly matched their tactic to their prey, suggesting that they learned how to hunt via repeated interactions. However, while the mean speed of predators was strongly and increasingly matched to the mean speed of the prey, further investigations revealed that this apparent increase was driven largely by one individual. Yet, predators also adjusted their degree of flexibility to their prey even if the change was not very strong (Figure 3 and Appendix S2: Table S1). In that regard, despite that only a portion of the population displayed large changes in specialization, and that adjustments to the prey were not as strong as we expected, our results still provide direct evidence that the mechanism underlying the resource-predictability hypothesis may indeed involve predators learning to adjust their degree of specialization based on their prey [@Weimerskirch.etal2005;@Weimerskirch2007;@Woo.etal2008;@Phillips.etal2017]. This is further supported by the fact that specialist and flexible hunters both achieved similar success throughout the study period.

An emerging pattern reported in predator-prey research is that predators and prey often match their phenotype to one another (i.e. positive trait covariances), which appears to be driven by an arms-race-like reciprocal phenotypic plasticity [@Kishida.etal2006;@Kishida.etal2009;@Edgell.Rochette2009a;@Mougi.etal2011;@McGhee.etal2013]. From an evolutionary perspective, @BrodieIII.BrodieJr.1999 showed in the garter snake (*Thamnophis sirtalis*) - roughskin newt (*Taricha granulosa*) system that such escalations may occur when prey develop defensive traits, which leads to stronger selection in predators, resulting in counteradaptations that circumvent the prey's defenses. In our system, the predator's success is negatively correlated with the prey's speed, which is an effective antipredator tactic [@FraserFranco.etal2022]. However, we found that as the covariance between predator speed and prey speed increased with experience, reminiscent of an arms race, the consequences of hunting faster prey decreased with experience. Moreover, this pattern was followed by a decrease in among individual variation in success with experience. This has important implications for the role of experience on predator-prey interactions, as our results suggest that learning enables predators to effectively develop countermeasures to the prey's defenses within their lifetime (i.e. at nonevolutionary timescales). It also provides empirical evidence for the hypothesis that reciprocal phenotypic plasticity through learning may attenuate selection exerted by prey on predators [@Anderson1995;@Ancel1999;@Borenstein.etal2006;@Paenke.etal2007].

@@

The interactions of predator and prey traits are probably the most important processes involved in population cycles and are expected to occur when predators match their phenotype to their prey [@abramsEvolutionPredatorPreyInteractions2000]. Yet, its has remained largely unknown whether this trait-matching results from predators learning how to capture their prey, in part because of the challenges of investigating direct interactions in the wild. By capitalizing on a virtual predator-prey system where we could directly monitor interactions, we found that experience drives individual variation in specialization, and that predators adjust their tactic to the behaviour of their prey. Moreover, individuals displayed large differences in the development of their expertise, which could result from differences in learning, as we found that almost half the population did not change their behaviour with experience. Our results have implications for predator-prey models of trait-matching, because such models are often studied at evolutionary timescales. Therefore, future studies should incorporate individual variation in experience to better predict under which ecological contexts specialization should be favoured over flexibility. Lastly, virtual systems are increasingly recognized among ecologists as effective systems to test hypotheses on consumer-resource interactions [@Barbe.etal2020;@Beauchamp2020;@Cere.etal2021;@FraserFranco.etal2022], but also in other fields such as citizen science and conservation [@Sandbrook.etal2015;@Redpath.etal2018;@Duthie.etal2021]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

>>>>

---
title: "Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame"
output: 
  word_document:
    reference_docx: Chapter2-styles.docx
    fig_caption: yes
    pandoc_args: ["-Fpandoc-crossref"] # to have eq. numbers
bibliography: references2.bib
csl: ecology.csl
---

@@

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing in different foraging tactics or on different resources irrespective of sexual, morphological, or age-related differences [@Estes.etal2003;@Tinker.etal2008;@Kernaleguen.etal2015;@Phillips.etal2017]. A growing body of evidence suggests that ecological interactions such as predator-prey interactions can drive this individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016]. When they hunt, predators often use techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their capacity to use them effectively is contingent on periods of extensive practice (i.e. experience). While hunting experience may be essential to maintain or improve foraging success, we have few empirical assessments of its role in predator foraging specialization [@Dukas2019] and its consequences on predator-prey interactions.

The development of foraging tactics is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. It is described as the relationship between success and cumulative experience, where extensive practice optimizes success when individuals reach an advanced level of experience. Empirical studies on human and non-human hunters show that experience optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. This body of work suggests that such optimization may thus promote individual foraging specialization (e.g. repeatedly using fast attacks) if the success of a hunting tactic is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that the costs/risks faced by hunters when switching foraging tactics may be offset through extensive practive, as individuals gain experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could result in both mechanisms operating at the same time within a predator population. For instance, some individuals may constantly encounter faster prey, which are proven to be more difficult to hunt in simulation and empirical studies [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. Thus, whether predators specialize or increase their flexibility with experience, learning to adjust their tactic to their prey should be essential to optimize their success.

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging [@Phillips.etal2017]. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specialization [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. This may in part be due to the limited information we have on the ecological contexts favouring specialization over flexibility in nature. However, in predator-prey systems, many studies have outlined that fluctuations in the predictability of prey encounters throughout a predator's lifetime may be a key factor shaping foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability (or resource-heterogeneity) hypothesis argues that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019]. When predator-prey interactions are directly observed, empirical evidence shows that predators often match, among other types of traits [see @Kishida.etal2006;@Hanifin.etal2008;@Brousseau.etal2018;@Reimche.etal2020], their locomotor and behavioural traits to those of their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022]. Yet, a recurring question that emerges from these studies is whether and how the behavioural adjustments of predators reflect individual differences in experience/learning [@Kelley.Magurran2011]. Thus, uncovering the role of learning in predator-prey systems where interactions are directly monitored would enable researchers to better predict the behavioural decisions and the success of predators when they are hunting.

@@

We tested whether foraging behaviour changed with experience, and compared the success of generalists vs specialists, using a multivariate, double-hierarchical generalized linear model (MDHGLM) [@Lee.Nelder2006;@Cleasby.etal2015;@westneatBiologyHiddenResidual2014;@Mitchell.etal2016a;@ODea.etal2022]. We categorized the predators by their experience based on the number of matches they played (i.e. their cumulative experience). Since we monitored all predator players throughout their experience, they appeared in all of the three following experience categories. First, we assigned predators the status of "novice" when they had a cumulative experience below 100 matches. For example, a predator player that played 500 matches in total would be labelled as a novice in its first 100 matches. Following this logic, we labelled predators as "intermediate" hunters when they had a cumulative experience between 100 and 299 and matches. Lastly, we defined predators as "advanced" hunters when they had a cumulative experience equal or above 300, with a maximum of 499 cumulated matches.

@@

At the individual level, we detected differences in the mean movement speed among predators (i.e. the intercept of the mean speed for the predator ID; Table 3). These differences remained stable from the novice to the intermediate stages, and then increased at the advanced stage to reach a level higher than when predators were novice (Table 3 and Appendix S1: Figure S4). Predators displayed considerable differences in their degree of hunting specialization, and these differences increased with experience (i.e. the intercept of sigma for the predator ID increases with experience; Table 3 and Appendix S1: Figure S4). Thus, as individual predators gain experience, the population becomes more heterogeneous, with an assemblage of both specialist and flexible hunters. Moreover, some individual predators switched from a flexible to a specialized hunting tactic, and vice-versa, as they gained experience (Figure 2). Importantly, 12.6% of the predator population displayed a large change in foraging specialization/flexibility with experience (i.e. $\ge$0.2 change in standard deviation), 44.7% displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation), while 42.7% remained relatively stable (i.e. $\le$0.05 change in standard deviation).

@@

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We selected the individuals displayed in the figure by first substracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selecting those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal or greater than 0.28 standard deviations."}

@@

Hunting experience is predicted to increase predator foraging specialization when resources are predictable [@Estes.etal2003;@Tinker.etal2008;@Tinker.etal2009]. However, when resources vary, others predict that learning with experience should increase behavioural flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013;@Wright.etal2022]. Our results show that there was no directional increase in either behavioural specialization or flexibility with experience at the population level. Instead, individual predators developed their own expertise and behavioural trajectory over successive encounters with prey; some specialized in fast-paced hunting, and others adopted a slower and more flexible tactic with experience. Specialized cursorial hunters encountered similar groups of prey while slower and flexible hunters encountered varying groups of prey, with both achieving similar success over the study period. With experience, predators increasingly matched their tactic to the speed of the prey that they encountered, which simultaneously relaxed the negative consequences of encountering faster prey. Together, our observations suggest that predators in *DBD* learn to adjust their hunting tactic to the behaviour of the prey that they encounter throughout their lifetime in the game. The resulting reciprocal behavioural escalation between the predators and the prey suggests that learning may be a catalyst for arms races [@Agrawal2001;@Kishida.etal2006;@Mougi.Kishida2009;@Mougi.etal2011].

@@

Empirical research in humans and animals shows that task proficiency often increases nonlinearly with experience and stabilizes at an expert level [reviewed in @Dukas2019]. However, we cannot derive conclusions at the individual-level from these studies either because their models assume the same increase in expertise for each individual, or because it is unclear whether their data includes repeated measures of individual success through time. Our analyses revealed that there was no population increase in expertise over time; instead, individual predators in *DBD* displayed distinct patterns of expertise acquisition (Figure 1, Appendix S1: Figure S1A-S1B). We predicted that this would occur if, by chance, predators differed in the prey that they encountered, because some may be easier and others more difficult to capture. In nature, prey can use a variety of defenses such as physical armaments, toxins, camouflage, or antipredator behaviours to escape predation [@BrodieIII.BrodieJr.1999;@Bowen.etal2002;@brodieCoEvolutionPredatorsPrey2010;@Nomura.etal2011;@Carey.Wahl2011]. In *DBD* we previously showed that faster speed and space coverage is an effective antipredator strategy used by the prey making them more difficult to capture [@FraserFranco.etal2022]. Other studies have also found that faster escapes are effective to escape predation [@Walker.etal2005;@Kelley.Magurran2011;Martin.etal2022]. Yet, despite that predators differed in the prey that they encountered, prey speed did not shape individual differences in expertise acquisition. It is however possible that other antipredator tactics mediated these individual responses. For example, it is well known in *DBD* that some prey groups use mobbing to confuse the predator or prevent it from capturing an injured individual. However, like natural predator-prey systems, direct predator-mobbing in *DBD* is dangerous and requires coordinated actions that are learned through experience [@Graw.Manser2007;@Dutour.etal2016;@Pitman.etal2017;@Dutour.etal2019;@Rampim.etal2020;@Szymkowiak2021], and was thus probably not a pervasive antipredator strategy. Other antipredator tactics known to confuse human and non-human predators during the attack sequence are unpredictable escapes and turning angles [@Jones.etal2011;@Briffa2013;@Herbert-Read.etal2017;@Moore.etal2017;@Richardson.etal2018]. While predators can adjust to unpredictable prey escapes [@Szopa-Comley.Ioannou2022], it could still be more arduous for a predator to learn and optimize its hunting success if it more often encountered prey using these tactics.

Alternatively, it is possible that individual variation in learning rates due to time delays between hunting events drove the observed differences in expertise acquisition, irrespective of the prey groups that predators encountered. For example, in *DBD*, a predator that played 300 matches in the span of six months might forget more critical information (e.g. prey escape patterns or muscle memory) than one that played 300 matches in the span of 6 days. While this has not been formally tested, a greater delay between exposure events is hypothesized to attenuate or even negate the relationship between success and experience [@Endler1991;@Wright.etal2022]. Moreover, the predators in *DBD* may have differed in their ability to detect prey. For example, it is widely demonstrated that predators learn visual cues such as coloration to recognize and avoid unpalatable prey [@Skelhorn.etal2016;@Ward-Fear.etal2020;@McLellan.etal2021], or habitat characteristics to locate food patches [@Ehlinger1989;@Hughes.etal1992;@Warburton2003]. @Szopa-Comley.etal2020 also found consistent among individual variation in the capacity to detect cryptic and conspicuous prey in three-spined sticklebacks (*Gasterosteus aculeatus*). Thus, individual predators should differ in their expertise if they do not learn equally how to exploit the visual cues that facilitate prey detection [@dukasEvolutionaryBiologyAnimal2004;@Dukas2019;@Wright.etal2022].

@@

The predator population displayed an increase in foraging flexibility when reaching the intermediate level of experience. As it reached an advanced level of experience, it returned to a level of flexibility that was similar to novice levels. This is commonly observed in juvenile predators across the animal kingdom, because exploring and learning different tactics at this developmental stage is crucial to become a skillful hunter [@Vehanen2003;@Johnson.Wilbrecht2011;@Thiers.etal2014;@deGrissac.etal2016]. Thus, predators in *DBD* were probably exploring and refining different tactics at this stage of experience. Within the population, individuals became increasingly distinct in their tactics as well as in their specialization with experience, which was related to the degree of variability in prey encounters. Predators that encountered groups of prey with similar speeds across matches specialized in hunting at high speeds, while those that encountered varying groups of prey adopted a slower and flexible strategy. Fast-paced hunting is indeed a highly specialized tactic suited for prey that use rapid escapes [@Endler1991;@Bro-Jorgensen2013;@Wilson.etal2018], while a slower and flexible tactic can be useful to minimize the consequences of uncertainty when hunting prey that use unpredictable escapes [@Szopa-Comley.Ioannou2022]. This is further supported by the fact that specialist and flexible hunters both achieved similar success throughout the study period. Interestingly, the behavioural response (i.e. tactic use and specialization) to experience by individual predators differed. For instance, 42.7% of the population did not alter their degree of specialization with experience (i.e. remaining either faster and more specialized, or slower and more flexible), while 57.3% of the population displayed a switch from a specialized to a flexible tactic or vice-versa with experience. Thus, our results provide direct evidence that the mechanism underlying the resource-predictability hypothesis may indeed involve predators learning at varying degrees how to specialize on their prey [@Weimerskirch.etal2005;@Weimerskirch2007;@Woo.etal2008;@Phillips.etal2017].

As predators gained experience, their tactic was increasingly matched to the speed of their prey. One interpretation could be that both predators and prey learned from each other in an arms-race-like reciprocal relationship [@Kishida.etal2006;@Kishida.etal2009;@Edgell.Rochette2009a;@Mougi.etal2011]. For instance, predators that encountered faster prey could have learned to increase their speed to better capture them, resulting in an increase in the speed of the prey to circumvent the predator's attacks. In this type of system, faster prey should be better at escaping both slower and faster predators [@Mougi.etal2011], which concords with what we previously observed in this sytem [@FraserFranco.etal2022]. @BrodieIII.BrodieJr.1999 argue that when prey express defensive phenotypes (e.g. physical armaments or venom), selection on predators should be strong because the consequences of being injured are greater than missing a meal. If predators improve their exploitative ability, then an arms-race should emerge, and the consequences of hunting dangerous prey should decrease [@BrodieIII.BrodieJr.1999]. Similarly, we found that has the predator gained experience, the negative consequences of encountering prey groups with higher speeds were relaxed, and the variation in success among predators decreased. However, in an arms-race, faster predators would have successfully preyed upon faster and slower prey and would have obtained higher success than slower predators [@Mougi.etal2011], which was not the case in the present study. A more likely interaction scenario is that predators and prey in *DBD* were involved in a matching response relationship [@Mougi.etal2011], because faster-specialist and slower-flexible preyed upon different prey but achieved equal success. Thus predators in *DBD* appear to optimize their success by learning to match their tactic to their prey with experience, suggesting that learning through reciprocal interactions may attenuate selection exerted by prey on predators [@Anderson1995;@Ancel1999;@Borenstein.etal2006;@Paenke.etal2007].

@@

The interactions of predator and prey traits are probably the most important processes involved in population cycles and should occur when predators match their phenotype to their prey [@abramsEvolutionPredatorPreyInteractions2000]. Yet, its has remained largely unknown whether this trait-matching results from predators learning how to capture their prey, in part because of the challenges of investigating direct interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that individual predators displayed large differences in the development of their expertise. With experience, half of the predator population changed their behaviour by developping either a specialized or flexible tactic that was matched to the speed of their prey. Our results could be useful to guide predator-prey models of trait-matching, which are often studied at evolutionary timescales. For instance, future studies could incorporate individual variation in experience in their models to better predict under which ecological/evolutionary contexts specialization should be favoured over flexibility. Lastly, virtual systems are increasingly recognized among ecologists as effective systems to test hypotheses on consumer-resource interactions [@Barbe.etal2020;@Beauchamp2020;@Cere.etal2021;@FraserFranco.etal2022], but also in other fields such as citizen science and conservation [@Sandbrook.etal2015;@Redpath.etal2018;@Duthie.etal2021]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
